Making Deep Q-learning Methods Robust to Time Discretization

### 2019-01-22 paper/icml_drau/advup.tex

\subsection{Scaling the Learning Rates}
\label{subsec:lr}
\input{pendulum_fig.tex}
For the discrete algorithm to admit a continuous time limit, the trajectories
of parameters must converge to well defined trajectories as $\deltat$ goes to
$0$.  The optimization steps should not be too large,
in which case parameter trajectories could jump in a single gradient step,
or too small, in which case the trajectories would become stationary as
$\deltat$ goes to $0$.  This in turn imposes conditions on the scalings of the
learning rates of algorithms robust to changes of the tine discretization.

The updates of $\theta$ and $\psi$ are described in
\eqref{eq:approx_bellman_A}.  To have parameter updates that do not explode or shrink to $0$ too
fast as $\deltat \rightarrow 0$, both $\delta \theta$ and
$\delta \psi$ should verify stochastic differential equations in the
limit, and notably,
have a deterministic component of order $\deltat$. As $\deltat$ goes to $0$,
a Taylor expansion yields (with $\deltas\deq s'-s$)
\begin{align}
	\delta Q_\deltat &= A(s, a)\,\deltat - r\,\deltat - \log(\gamma)
	\,V(s)\,\deltat - \partial_s V(s) \,\deltas\nonumber \\&- \frac{(\deltas)^T \partial_{s}^2 V(s) \deltas}{2} + \smallo\left(\deltat\right)\\
	\delta \theta_\deltat &= \eta^V_\deltat \partial_{\theta} V(s)
	\frac{\delta Q_\deltat}{\deltat} \\%+ \smallo\left(\eta^V_\deltat\right)\\
	\delta \psi_\deltat &= \eta^A_\deltat \partial_{\psi} A(s)
	\frac{\delta Q_\deltat}{\deltat} %+ \smallo\left(\eta^A_\deltat\right).
\end{align}

If we set 
$\eta^V_\deltat = \alpha^V \deltat$ and $\eta^A_\deltat = \alpha^A
\deltat$, the parameter updates will follow a Euler discretization scheme
for a continuous-time equation, thus presumably converging. This is
actually the only possibility; more precisely,

@@

\TODO{Thus, the right learning rate scaling } results in a fullly specified algorithm,
\emph{Deep Advantage Updating} (DAU).
Pseudocode for the discrete actions algorithm is given in Alg.~\ref{alg:dau}.
Pseudocode for continuous actions is given in the supplementary
material.

>>>>

\subsection{Scaling the Learning Rates}
\label{subsec:lr}
\input{pendulum_fig.tex}
For the algorithm to admit a continuous-time limit, the discrete-time trajectories
of parameters must converge to well defined trajectories as $\deltat$ goes to
$0$.
% The optimization steps should not be too large,
% in which case parameter trajectories could jump in a single gradient step,
% or too small, in which case the trajectories would become stationary as
% $\deltat$ goes to $0$.
This in turn imposes conditions on the scalings of the
learning rates of algorithms robust to changes of the time discretization.

Informally, in the updates of $\theta$ and $\psi$, the quantity $\delta
Q_\deltat$ is of order $O(\deltat)$, because $s'=s+O(\deltat)$ in a
near-continuous system. Therefore, $\delta
Q_\deltat/\deltat$ is of order $O(1)$, so that the gradients used to
update $\theta$ and $\psi$ are of order $O(1)$. Therefore, if the
learning rates are of order $\deltat$, one would expect 
the parameters $\theta$ and $\psi$ to move by $O(\deltat)$ in
each time interval $\deltat$, thus hopefully converging to smooth
continuous-time trajectories.

% The updates of $\theta$ and $\psi$ are described in
% \eqref{eq:approx_bellman_A}.  To have parameter updates that do not explode or shrink to $0$ too
% fast as $\deltat \rightarrow 0$, both $\delta \theta$ and
% $\delta \psi$ should verify stochastic differential equations in the
% limit, and notably,
% have a deterministic component of order $\deltat$. As $\deltat$ goes to $0$,
% a Taylor expansion yields (with $\deltas\deq s'-s$)
% \begin{align}
% 	\delta Q_\deltat &= A(s, a)\,\deltat - r\,\deltat - \log(\gamma)
% 	\,V(s)\,\deltat - \partial_s V(s) \,\deltas\nonumber \\&- \frac{(\deltas)^T \partial_{s}^2 V(s) \deltas}{2} + \smallo\left(\deltat\right)\\
% 	\delta \theta_\deltat &= \eta^V_\deltat \partial_{\theta} V(s)
% 	\frac{\delta Q_\deltat}{\deltat} \\%+ \smallo\left(\eta^V_\deltat\right)\\
% 	\delta \psi_\deltat &= \eta^A_\deltat \partial_{\psi} A(s)
% 	\frac{\delta Q_\deltat}{\deltat} %+ \smallo\left(\eta^A_\deltat\right).
% \end{align}

The next theorem confirms this analysis: if we set 
$\eta^V_\deltat = \alpha^V \deltat$ and $\eta^A_\deltat = \alpha^A
\deltat$, the parameter updates follow a Euler discretization scheme
for a continuous-time equation and have a well-defined limit behavior. On
the other hand, with smaller learning rates the parameters do not move,
and can explode instantaneously with larger learning rates.

@@

\TODO{Thus, the right learning rate scaling } results in a fully specified algorithm,
\emph{Deep Advantage Updating} (DAU).
Pseudocode for the discrete actions algorithm is given in Alg.~\ref{alg:dau}.
Pseudocode for continuous actions is given in the supplementary
material.
